---
title: "MetabolicsBiomechanics_2026-05-19"
author: "Caitlin Banks"
date: "2024-05-14"
output: 
  html_document:
  toc: true
  toc_float: true
  theme: flatly
  highlight: tango
  df_print: paged
  code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Load Packages

```{r, echo=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(svglite)
library(rstatix)
```

## Load Data
```{r}
netpower <- read.csv("[filename]/SummaryData_NetPower_2025-03-20.csv",header=TRUE)

biomechanics <- read.csv("[filename]/DynamicTread_KineticsKinematics_CBcontrols_2025-04-17.csv",header=TRUE)

asymmetry <- read.csv("[filename]/DynamicTread_Asymmetries_CBcontrols_2025-04-17.csv",header=TRUE)

biomechanics$Time <- factor(biomechanics$Time, levels=c('first30','last30'))

cross_correlation <- read_excel("[filename]/DynamicTread_CBControls_CrossCorrelation_Averages_2026-03-03.xlsx",sheet='Last30_AvgCrossCorrelation')

delta_netpower <- read.csv("[filename]/ChangeData_NetPower_2026-05-15.csv",header=TRUE)
```

## Filter Down to Dynamic Conditions
No open loop
No baselines
Last 30 strides - first 30 is not a good comparison to net power because net power is measured from last 2 minutes of trial
Removing GR0332 and GR0333 for incomplete data/exclusion
```{r}
biomech_data <- biomechanics %>% filter(Time == 'last30' & Condition != 'Open Loop - Pre' & Condition != 'Open Loop - Post')

biomech_data$Condition <- factor(biomech_data$Condition, levels=c('Fast','Slow','Accelerate','Decelerate','Baseline075','Baseline1125','Baseline150'))

biomech_data_R <- biomech_data %>% filter(Leg == 'Right')
biomech_data_L <- biomech_data %>% filter(Leg == 'Left')

power_data <- netpower %>% filter(Event!='Open Loop - Pre' & Event != 'Open Loop - Post' & Event != 'Baseline 0.75 m/s' & Event != 'Baseline 1.125 m/s' & Event != 'Baseline 1.5 m/s' & Event != 'Post 1.5 m/s')
power_data <- power_data %>% filter(Subject != "GR0332" & Subject != "GR0333")

power_data$Event <- factor(power_data$Event, levels=c('Fast','Slow','Accelerate','Decelerate'))

asymmetry_data <- asymmetry %>% filter(Trial != 'Baseline075' & Trial != 'Baseline1125' & Trial != 'Baseline150')
asymmetry_data <- asymmetry %>% filter(Subjects != "GR0332" & Subjects != "GR0333")

asymmetry_data$Condition <- factor(asymmetry_data$Trial, levels=c('Fast','Slow','Accelerate','Decelerate'))

delta_netpower <- delta_netpower %>% filter(Event != 'Baseline 1.125 m/s')
delta_netpower <- delta_netpower %>% filter(Subject != 'GR0332' & Subject != 'GR0333')
```

```{r Join Tables}
all <- left_join(power_data,biomech_data,by=c('Subject'='Subjects','Event'='Condition'))
#all$Dynamic.Trial = NULL

all_asym <- left_join(power_data,asymmetry_data,by=c('Subject'='Subjects','Event'='Condition'))

delta_asym <- left_join(delta_netpower,asymmetry_data,by=c('Subject'='Subjects','Event'='Condition'))
```

```{r Kinematic Plots - Data Values Colored by Leg}
ggplot(all, aes(x=`Net.Power`,y=StepLength_Norm,color=Leg))+geom_point()+facet_wrap(~Event)+labs(title="Normalized Step Length")
#ggsave("S:/Banks/DynamicTread/Metabolic Data/Figures/SL_Norm_Correlations_.png",dpi=300)

ggplot(all, aes(x=`Net.Power`,y=StepTime,color=Leg))+geom_point()+facet_wrap(~Event)+labs(title="Step Time (s)")

ggplot(all, aes(x=`Net.Power`,y=LeadingLimbAngle,color=Leg))+geom_point()+facet_wrap(~Event)+labs(title="Leading Limb Angle (deg)")

ggplot(all, aes(x=`Net.Power`,y=TrailingLimbAngle,color=Leg))+geom_point()+facet_wrap(~Event)+labs(title="Trailing Limb Angle (deg)")
```
```{r Kinetic Plots}
ggplot(all, aes(x=`Net.Power`,y=PeakPropulsion,color=Leg))+geom_point()+facet_wrap(~Event)+labs(title="Peak Propulsion")
#ggsave("[filepath]/SL_Norm_Correlations_.png",dpi=300)

ggplot(all, aes(x=`Net.Power`,y=PeakBraking,color=Leg))+geom_point()+facet_wrap(~Event)+labs(title="Peak Braking")

ggplot(all, aes(x=`Net.Power`,y=PropulsionImpulse,color=Leg))+geom_point()+facet_wrap(~Event)+labs(title="Propulsion Impulse")

ggplot(all, aes(x=`Net.Power`,y=BrakingImpulse,color=Leg))+geom_point()+facet_wrap(~Event)+labs(title="Braking Impulse")
```


## Building a Regression Model
```{r}
lm_PeakPropulsion <- lm(`Net.Power` ~ PeakPropulsion + Leg + Event, data=all)
summary(lm_PeakPropulsion)
```

## Asymmetries
```{r Visualizing Correlations}
all_asym$Event <- factor(all_asym$Event, levels=c("Fast", "Slow", "Accelerate","Decelerate"))

ggplot(all_asym, aes(y=`Net.Power`,x=StepLengthAsymmetry))+
  geom_point(size=3)+
  facet_wrap(~Event)+
  labs(x="Step Length Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

ggplot(all_asym, aes(y=`Net.Power`,x=StepTimeAsymmetry))+
  geom_point(size=3)+
  facet_wrap(~Event)+
  labs(x="Step Time Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

ggplot(all_asym, aes(y=`Net.Power`,x=LeadingLimbAngleAsymmetry))+
  geom_point(size=3)+
  facet_wrap(~Event)+
  labs(x="Leading Limb Angle Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

ggplot(all_asym, aes(y=`Net.Power`,x=TrailingLimbAngleAsymmetry))+
  geom_point(size=3)+
  facet_wrap(~Event)+
  labs(x="Trailing Limb Angle Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

ggplot(all_asym, aes(y=`Net.Power`,x=PeakPropulsionAsymmetry))+
  geom_point(size=3)+
  facet_wrap(~Event)+
  labs(x="Peak Propulsion Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

ggplot(all_asym, aes(y=`Net.Power`,x=PeakBrakingAsymmetry))+
  geom_point(size=3)+
  facet_wrap(~Event)+
  labs(x="Peak Braking Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

ggplot(all_asym, aes(y=`Net.Power`,x=PropulsionImpulseAsymmetry))+
  geom_point(size=3)+
  facet_wrap(~Event)+
  labs(x="Propulsion Impulse Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

ggplot(all_asym, aes(y=`Net.Power`,x=BrakingImpulseAsymmetry))+
  geom_point(size=3)+
  facet_wrap(~Event)+
  labs(x="Braking Impulse Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))
```
# Modeling Asymmetries vs. Net Power

## Step Length Asymmetry
```{r SL Asym vs Net Power}
lm_SLasym <- lm(`Net.Power` ~ StepLengthAsymmetry + Event, data=all_asym)
summary(lm_SLasym)

ggplot(all_asym, aes(y=`Net.Power`,x=StepLengthAsymmetry,color=Event))+
  geom_point(size=3)+
  scale_color_manual(values=c("#ff00ff","#40a6ff","#ff8000","#ffbf00"))+
  labs(x="Step Length Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

#ggsave("[filepath]/NetPowervsSLA_2025-05-14.svg",width=7,height=4,units="in",dpi=600)

fast_asym <- all_asym %>% filter(Event=="Fast")
lm_SLasym_fast <- lm(`Net.Power` ~ StepLengthAsymmetry,data=fast_asym)
summary(lm_SLasym_fast)

cor.test(fast_asym$`Net.Power`,fast_asym$StepLengthAsymmetry)
```
### Step Time Asymmetry
```{r ST Asym vs Net Power}
lm_STasym <- lm(`Net.Power` ~ StepTimeAsymmetry + Event, data=all_asym)
summary(lm_STasym)
```
## Leading Limb Angle Asymmetry
```{r LLA Asym vs Net Power}
lm_LLAasym <- lm(`Net.Power` ~ LeadingLimbAngleAsymmetry + Event, data=all_asym)
summary(lm_LLAasym)
```
## Trailing Limb Angle Asymmetry
```{r TLA Asym vs Net Power}
lm_TLAasym <- lm(`Net.Power` ~ TrailingLimbAngleAsymmetry + Event, data=all_asym)
summary(lm_TLAasym)

ggplot(all_asym, aes(y=`Net.Power`,x=TrailingLimbAngleAsymmetry,color=Event))+
  geom_point(size=3)+
  scale_color_manual(values=c("#ff00ff","#40a6ff","#ff8000","#ffbf00"))+
  labs(x="Trailing Limb Angle Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

#ggsave("[filepath]/NetPowervsTLAA_2025-05-14.svg",width=7,height=4,units="in",dpi=600)
```

## Peak Propulsion Asymmetry
```{r PP Asym vs Net Power}
lm_PPasym <- lm(`Net.Power` ~ PeakPropulsionAsymmetry + Event, data=all_asym)
summary(lm_PPasym)

ggplot(all_asym, aes(y=`Net.Power`,x=PeakPropulsionAsymmetry,color=Event))+
  geom_point(size=3)+
  scale_color_manual(values=c("#ff00ff","#40a6ff","#ff8000","#ffbf00"))+
  labs(x="Peak Propulsion Asymmetry",y="Net Power (W/kg)")+
  geom_vline(xintercept=0,linetype="dashed")+
  theme_classic() +
  theme(text=element_text(size=16))

#ggsave("S:/Banks/DynamicTread/Metabolic Data/SummaryData_Control/Figures/NetPowervsPPA_2025-05-14.svg",width=7,height=4,units="in",dpi=600)
```
## Peak Braking Asymmetry
```{r PB Asym vs Net Power}
lm_PBasym <- lm(`Net.Power` ~ PeakBrakingAsymmetry + Event, data=all_asym)
summary(lm_PBasym)
```
## Propulsion Impulse Asymmetry
```{r PI Asym vs Net Power}
lm_PIasym <- lm(`Net.Power` ~ PropulsionImpulseAsymmetry + Event, data=all_asym)
summary(lm_PIasym)
```
## Braking Impulse Asymmetry
```{r BI Asym vs Net Power}
lm_BIasym <- lm(`Net.Power` ~ BrakingImpulseAsymmetry + Event, data=all_asym)
summary(lm_BIasym)
```
# Looking at abs(asymmetry) and change in power from 1.125 condition
```{r}
# Calculate abs(asymmetry) values for step length, trailing limb angle, and peak propulsion asymmetries
delta_asym$abs_StepLengthAsymmetry <- abs(delta_asym$StepLengthAsymmetry)
delta_asym$abs_TrailingLimbAngleAsymmetry <- abs(delta_asym$TrailingLimbAngleAsymmetry)
delta_asym$abs_PeakPropulsionAsymmetry <- abs(delta_asym$PeakPropulsionAsymmetry)
```

## Correlation Plots
```{r Step Length Asymmetry and Delta Net Power}
lm_SLasym_delta_abs <- lm(`Delta.NetPower.` ~ abs_StepLengthAsymmetry + Event, data=delta_asym)
summary(lm_SLasym_delta_abs)

ggplot(delta_asym, aes(y=`Delta.NetPower.`,x=abs_StepLengthAsymmetry,color=Event))+
  geom_point(size=3)+
  scale_color_manual(values=c("#ff00ff","#40a6ff","#ff8000","#ffbf00"))+
  labs(x="Absolute Step Length Asymmetry",y="Change in Net Power (W/kg)")+
  theme_classic() +
  theme(text=element_text(size=16))

#ggsave("[filepath]/DeltaNetPowervsSLA_2026-05-15.svg",width=7,height=4,units="in",dpi=600)

```

```{r TLA Asymmetry and Delta Net Power}
lm_TLAasym_delta_abs <- lm(`Delta.NetPower.` ~ abs_TrailingLimbAngleAsymmetry + Event, data=delta_asym)
summary(lm_TLAasym_delta_abs)

ggplot(delta_asym, aes(y=`Delta.NetPower.`,x=abs_TrailingLimbAngleAsymmetry,color=Event))+
  geom_point(size=3)+
  scale_color_manual(values=c("#ff00ff","#40a6ff","#ff8000","#ffbf00"))+
  labs(x="Absolute Trailing Limb Angle Asymmetry",y="Change in Net Power (W/kg)")+
  theme_classic() +
  theme(text=element_text(size=16))

#ggsave("[filepath]/DeltaNetPowervsTLA_2026-05-15.svg",width=7,height=4,units="in",dpi=600)
```

```{r Propulsion Asymmetry and Delta Net Power}
lm_PPasym_delta_abs <- lm(`Delta.NetPower.` ~ abs_PeakPropulsionAsymmetry + Event, data=delta_asym)
summary(lm_PPasym_delta_abs)

ggplot(delta_asym, aes(y=`Delta.NetPower.`,x=abs_PeakPropulsionAsymmetry,color=Event))+
  geom_point(size=3)+
  scale_color_manual(values=c("#ff00ff","#40a6ff","#ff8000","#ffbf00"))+
  labs(x="Absolute Peak Propulsion Asymmetry",y="Change in Net Power (W/kg)")+
  theme_classic() +
  theme(text=element_text(size=16))

#ggsave("[filepath]/DeltaNetPowervsPPA_2026-05-15.svg",width=7,height=4,units="in",dpi=600)
```


# Cross Correlation Data for Speed Profile Matching
```{r Cross Correlations}
cross_correlation <- cross_correlation %>% filter(Subject!='Average')

cross_correlation_long <- cross_correlation %>% pivot_longer(-Subject)

cross_correlation_long$value[is.nan(cross_correlation_long$value)] <- NA
cross_correlation_long$value <- as.numeric(cross_correlation_long$value)

res.aov <- anova_test(data=cross_correlation_long, dv = value, wid=Subject, within = name)
get_anova_table(res.aov)
```