---
title: "MetabolicSummary_2025-05-19"
author: "Caitlin Banks"
date: "2022-09-28"
output: 
  html_document:
  toc: true
  toc_float: true
  theme: flatly
  highlight: tango
  df_print: paged
  code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Setup
## Load Packages

```{r, warning=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
library(plyr)
library(dplyr)
library(svglite)
library(rstatix)
library(DescTools)
library(ggiraph)
library(RColorBrewer)
```
## Load Data

```{r}
netpower <- read_excel("[filepath]/SummaryData_NetPower_17May2024.xlsx",sheet='NetPower')

netpower$Event <- factor(netpower$Event, levels=c('Baseline 0.75 m/s','Baseline 1.125 m/s','Baseline 1.5 m/s','Open Loop - Pre','Fast','Slow','Accelerate','Decelerate','Open Loop - Post','Post 1.5 m/s'))
```
Excluding GR0332 and GR0333 because of missing data
  -GR0332 blocks 7-10 because the metabolic data are not trustworthy after removing and adjusting the mask
  -GR0333 does not have a Post150 due to technical issues with metabolic cart computer & participant needing to leave

```{r Complete Cases}
netpower_complete <- netpower %>% filter(Subject!='GR0332' & Subject!='GR0333')
```

```{r Set Color Palettes}
colorCount <- length(unique(netpower$Subject))
getPalette <- colorRampPalette(brewer.pal(9,"Set1"))
```

# Plot data
## Individual Subjects Plus Average

```{r}
netpower$`Net Power` <- as.numeric(as.vector(netpower$`Net Power`))

# Colored by individual
ggplot(data=netpower_complete, aes(x=Event, y=`Net Power`, group=Subject, color=Subject, na.rm=TRUE)) + theme_minimal() + theme(axis.text.x=element_text(angle=45)) + ylim(c(0,4.5)) + labs(x=NULL,y='Net Power (W/kg)') + stat_summary(aes(y=`Net Power`,group=Subject),fun.data=mean_se,color='grey70',fill='grey70',geom='ribbon',group=1,linewidth=1) + stat_summary(aes(y=`Net Power`, group=Subject),fun=mean,color='black',geom='line',group=1,linewidth=1)+ geom_line() + geom_point() + scale_fill_manual(values=getPalette(colorCount))


# All individuals grey, mean +/- SD error bar
ggplot(data=netpower_complete, aes(x=Event, y=`Net Power`, group=Subject, na.rm=TRUE)) + theme_minimal() + theme(axis.text.x=element_text(angle=45)) + ylim(c(0,4.5)) + labs(x=NULL,y='Net Power (W/kg)') + 
  stat_summary(aes(y=`Net Power`, group=Subject), fun.data=function(x) c(y = mean(x), ymin = mean(x) - sd(x), ymax = mean(x) + sd(x)), color='black', geom='errorbar',group=1,width=0.5,size=1) + 
  stat_summary(aes(y=`Net Power`, group=Subject),fun=mean,color='black',geom='line',group=1,size=1)+ stat_summary(aes(y=`Net Power`, group=Subject),fun=mean,color='black',geom='point',group=1,size=2) +
  geom_line(color='grey70') + geom_point(size=1,color='grey70') 

#ggsave("[filepath]SummaryData_NetPower_20Sept2024_grayscale.svg",width=7,height=4,units="in",dpi=300)
```
# Stats and Calculations
## Calculate Mean and Standard Deviation
```{r}
# Function to calculate the mean and standard deviation for each group
data_summary <- function(data,varname,groupnames){
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum <- ddply(data, groupnames, .fun=summary_func, varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}

netpower_summary <- data_summary(netpower_complete,varname="Net Power",groupnames = "Event")

netpower_summary$Subject <- "Average"
```

## Some Simple Stats
```{r}
identify_outliers(data.frame(netpower_complete$`Net Power`))

shapiro.test(netpower_complete$`Net Power`)

bartlett.test(`Net Power`~Event,data=netpower_complete)
```
There is 1 outlier point, not considered extreme (+/-3IQR).
After reviewing the data, I believe all of them.

Shapiro-Wilk passes, data are normally distributed

Bartlett's test passes, we have homogeneity of variance


```{r Repeated Measures ANOVA}
res.aov <- anova_test(data=netpower_complete, dv = `Net Power`, wid= Subject, within = Event)
get_anova_table(res.aov)
```
There's a significant effect of Event

```{r Post-hoc tests - Dunnetts}
# Remove Baseline 0.75 and 1.125 conditions, only testing 1.5 against DT's
netpower_complete_dunnett <- netpower_complete %>% filter(Event!="Baseline 0.75 m/s" & Event!="Baseline 1.5 m/s" & Event!="Post 1.5 m/s")
DunnettTest(netpower_complete_dunnett$`Net Power`,g=netpower_complete_dunnett$Event,control="Baseline 1.125 m/s")

# Playing around - is DT different from 1.5?
netpower_complete_dunnett_1.5 <- netpower_complete %>% filter(Event!="Baseline 0.75 m/s" & Event!="Baseline 1.125 m/s" & Event!="Post 1.5 m/s")
DunnettTest(netpower_complete_dunnett_1.5$`Net Power`,g=netpower_complete_dunnett_1.5$Event,control="Baseline 1.5 m/s")

# Playing around - is DT different from post 1.5?
netpower_complete_dunnett_post <- netpower_complete %>% filter(Event!="Baseline 0.75 m/s" & Event!="Baseline 1.125 m/s" & Event!="Baseline 1.5 m/s")
DunnettTest(netpower_complete_dunnett_post$`Net Power`,g=netpower_complete_dunnett_post$Event,control="Post 1.5 m/s")
```

## Baseline Speed Scaling
```{r Baseline Speeds Only}
baselines <- netpower_complete %>% filter(Event=="Baseline 0.75 m/s"|Event=="Baseline 1.125 m/s"|Event=="Baseline 1.5 m/s")

res.aov3 <- anova_test(data=baselines, dv = `Net Power`, wid= Subject, within = Event)
get_anova_table(res.aov3)
```
```{r Baseline Speeds Pairwise}
baselines$Subject <- factor(baselines$Subject)

pwc3 <- pairwise.t.test(baselines$`Net Power`,baselines$Event, p.adjust.method='bonferroni',paired=TRUE)
pwc3$p.value
```

## Pre-Post 1.5 t-test
```{r Normality Test on Sub-Sample}
prepost <- netpower_complete %>% filter(Event=="Baseline 1.5 m/s"|Event=="Post 1.5 m/s")
differences <- with(prepost,`Net Power`[Event=="Baseline 1.5 m/s"] - `Net Power`[Event=="Post 1.5 m/s"])
shapiro.test(differences)
```

```{r Paired T-test}
t.test(differences,mu=0)
```
Technically I'm doing a one-sample t-test of the differences between the groups, which should mathematically be the same thing as a paired t-test

## Pre-post open loop test
```{r Normality Test on Open Loop Sub-Sample}
prepost_openloop <- netpower_complete %>% filter(Event=="Open Loop - Pre"|Event=="Open Loop - Post")
differences_openloop <- with(prepost_openloop,`Net Power`[Event=="Open Loop - Pre"] - `Net Power`[Event=="Open Loop - Post"])
shapiro.test(differences_openloop)
t.test(differences_openloop,mu=0)
```

## Order Effects
```{r}
netpower_complete$Block <- factor(netpower_complete$Block)

ggplot(data=netpower_complete, aes(x=Block, y=`Net Power`, group=Subject, color=Subject, na.rm=TRUE)) + theme_minimal() +  ylim(c(0,4.5)) + labs(x=NULL,y='Net Power (W/kg)') + stat_summary(aes(y=`Net Power`,group=Subject),fun.data=mean_se,color='grey70',fill='grey70',geom='ribbon',group=1,linewidth=1) + stat_summary(aes(y=`Net Power`, group=Subject),fun=mean,color='black',geom='line',group=1,linewidth=1)+ geom_line() + geom_point() + scale_color_brewer(palette="Paired")

# ANOVA of all blocks
res.aov_blockorder <- anova_test(data=netpower_complete, dv = `Net Power`, wid= Subject, within = Block)
get_anova_table(res.aov_blockorder)

netpower_complete_dunnett <- netpower_complete %>% filter(Block!="1" & Block!="3" & Block!="10")
DunnettTest(netpower_complete_dunnett$`Net Power`,g=netpower_complete_dunnett$Block,control="2")

netpower_complete_metabolic_dunnett <- netpower_complete %>% filter(Block!="1" & Block!="2" & Block!="3" & Block!="10")
DunnettTest(netpower_complete_metabolic_dunnett$`Net Power`,g=netpower_complete_metabolic_dunnett$Block,control="5")

# Block order with mean +/- SD errorbar
ggplot(data=netpower_complete, aes(x=Block, y=`Net Power`, group=Subject, na.rm=TRUE)) + theme_minimal() + ylim(c(0,4.5)) + labs(x='Testing Block',y='Net Power (W/kg)') +  stat_summary(aes(y=`Net Power`,group=Subject), fun.data = function(x) c(y = mean(x), ymin = mean(x) - sd(x), ymax = mean(x) + sd(x)), color='black',geom='errorbar',group=1,width=0.5, size=1) + stat_summary(aes(y=`Net Power`, group=Subject),fun=mean,color='black',geom='line',group=1,linewidth=1)+ stat_summary(aes(y=`Net Power`, group=Subject),fun=mean,color='black',geom='point',group=1,linewidth=2) + geom_line(color='grey70') + geom_point(linewidth=1,color='grey70')

#ggsave("[filepath]OrderEffects_NetPower_2026-05-12.svg",width=7,height=4,units="in",dpi=600)
```
```{r Order Effects Pairwise}
pwc_order <- netpower_complete %>% pairwise_t_test(`Net Power`~Block, paired=TRUE, p.adjust.method = "bonf")
pwc_order
```
No notable differences between dynamic tread blocks
